Taking Dyads Seriously

### 2018-02-14 paper/4_app.tex
some tests

An important takeaway here is that many scholars are forced to make knowledge claims based on the statistical significance of a small set of covariates, or the differences between these covariates. These differences may change dramatically when interdependencies are taken into account directly, as they do in the case of this study. This outcome follows from AME's ability to better account for the dependencies discussed in the previous section, whereas GLM approaches explicitly assume observational independence conditional on the specified covariates. As this is a widely-known limitation of GLM approaches, scholars often attempt to account for clustering of observations by including additional variables and adjusting the standard errors of the resulting estimates. At best, this method introduces noise and imprecision into results, and at worst can produce misleading outcomes. Beyond just comparing parameter estimates, we examine how well each approach can represent the data generating process using an out-of-sample cross validation strategy. Specifically, for each study, we randomly divide the data into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.

@@

The result of this procedure is a set of sociomatrices $\bm \hat Y$, in which each entry $\hat y_{ij,t}$ is a predicted value obtained from using a subset of the data that does not include $y_{ij,t}$.\footnote{For more details on this type of cross-validation strategy see \citet{minhas:etal:2016:arxiv}.} We summarize the performance of the various models in Table~\ref{tab:modelPerfSumm} below. For the binary models we  provide the area under the Receiver Operator Characteristic (ROC) and Precision Recall (PR) curves. Only one of the studies here had a continuous dependent variable and for this we provide the root mean squared error (RMSE) and root median squared error (RMDSE).\footnote{More details on the performance of each of these models can be found in the Appendix.} For each of the replications, we find that the AME approach substantially outperforms the original models in terms of out-of-sample predictive performance.

@@

In the original model, the variable ``Pers/Democ Directed Dyad" (which represents a Personalist $\rightarrow$ Democractic directed dyad) has a positive effect while the variable ``Democ/Personalist Directed Dyad'' is too imprecisely measured to indicate a direction. In our re-estimation using the AME framework, we also find that Pers-Democ directed dyad has a positive effect while Democ-Pers directed dyad is still too imprecisely measured to indicate a direction. Using this model, however, we can no longer conclusively say that the Pers/Democratic coefficient is larger than the Democ/Personalist one. Our re-estimation using the AME approach therefore cast some doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies, though many of the major effects  uncovered by the authors remains robust when accounting for dyadic interdependencies. %Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework suggesting that some of their effects were actually the products of interdependencies.

@@

\citet{mcdonald:2004} studied whether trade promotes peace between nations. \citet[p. 547]{mcdonald:2004} introduced the argument that free trade between states ``makes conflict less likely because of its efficiency over conquest in acquiring resources\ldots''. Accordingly, he provided evidence challenging the generalized linkage between peace and trade and refined the measurement of the standard ``trade'' variable, arguing that \textit{free} trade, rather than trade alone, reduces the likelihood of conflict between states. His key hypothesis is that greater levels of protection increase the probability of interstate conflict, an argument that builds on the work of classic liberalism and connects free trade to the power of domestic audiences. \citet{mcdonald:2004} measured free trade in two ways. The first, the ``protection'' variable measures the proportion of customs revenue divided by total imports in the state that possesses the greater such ratio in each dyad. reflects the notion that larger protected sectors generate greater societal pressures resulting in pockets of support for war. This measure thus captures the score of the state in the dyad that possesses higher barriers to trade. \citet[p. 560]{mcdonald:2004} also includes a measure of economic integration  calculated as ``the lower proportion of total dyadic trade (imports plus exports) divided by state $i$'s GDP or total dyadic trade divided by state $j$'s GDP''. The (binary) dependent variable is the onset of a new militarized interstate dispute within a given dyad.  \citet{mcdonald:2004} employs logistic regression to examine the  putative statistical significance of these variables. The models  include splines to correct for temporal dependence, and robust  standard errors clustered on each dyad.

@@

When we compare the results of Rose's original OLS model to the AME model accounting for network dependencies, the results are similar. The main result of the model---the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables---does not prevail in the AME model. However, in a sense the AME result provides even stronger evidence against a positive effect of WTO membership: in the original model, this negative linkage is not significant, but in the AME estimation this negative linkage is more precisely estimated. The fact that this variable attains conventional statistical significance under an AME estimation shows an additional value of this approach to applied researchers -- while accounting for interdependencies and removing that source of bias will remove false positive findings, it can also avoid false negative findings where the bias prevented a researcher from uncovering a consistent relationship.\footnote{This benefit comes up less often in these replications because most published studies are positive, rather than null findings.} In both cases we find  a rejection of the conventional wisdom, that WTO membership increases trade. While in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model. Of course, the controversial result of the paper, strengthened in the AME analysis might be an artifact of measurement error. \citep{tomz:etal:2007}. While in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.

\begin{figure}
	\includegraphics[width=\textwidth]{rose_aeff_top10.pdf}
	\caption{Nodal Random Effects for AME estimation of Rose (2004), for the countries with the highest and lowest Random Effects.}\label{fig:roser}
\end{figure}
\FloatBarrier

@@

\citet{weeks:2012} examines the influence of domestic institutions on the initiation of military conflicts by autocratic leaders.  She argues that in some circumstances autocrats are held accountable for their foreign policy decisions. She adds the nuance that autocratic audiences are not homogeneous. When the autocratic regime is nonmilitary, domestic audiences do not favor military actions, but in military autocracies this is not the case. Further she argues that in personalistic regimes without a military or civilian domestic audience, the leaders tend to be more likely to employ military force in their foreign policy.  To study this question, she uses a dyadic design in which the dependent variable is ``whether country A in a directed dyad initiated military conflict against country B during year t'' (page 337).  One major innovation in her study resides in the nuanced way in which she conceptualizes and codes regime type into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen. She also includes a variety of putative control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.'' She uses a logistic regression, but follows \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with fixed, unit effects. The analysis is done for dyads, but is considered to be from the perspective of the actor that initiated the dispute. The basic finding is that a) juntas, boss type, and strongmen type regimes are more likely to initiate conflict than machine-type regimes (and maybe democracies) and that b) machine-type regimes are no more belligerent than democracies.   These insights are mainly emphasized in the paper by the parameter estimates depicted in Tables 1 and 2 (pages 339--340) from the paper. She argued that ignoring important nuances between different types of autocracies hinders our understanding of the initiation of military conflict by autocracies.

\begin{figure}[!h]
	\includegraphics[width=\textwidth]{weeks_circPlot.png}
	\caption{\label{fig:weekscirc} Visualization of multiplicative effects for Weeks (2012). Each circle designates a country and the color corresponds to the legend at the center of the visualization. Countries that cluster together in the outer ring are those that were found by the model to have similar sending patterns, meaning that they tend to send conflict to similar sets of countries. The inner ring clusters countries by the similarity of who they receive conflict from. 
	%Blue represents groups with common sending patterns and red represents groups with common receiving patterns.
	}
\end{figure}
\FloatBarrier

The re-estimation of \citet{weeks:2012} has the sharpest divergence between the GLM results and those of the AME Model. In Weeks' initial models,  she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses, and Strong-men are more conflict-prone. When we examine the AME results, we find that none of these values are distinguishable from zero. Similarly, we find less pronounced effects for military capabilities. One explanation for this divergence is the AME model's ability to account for third-order effects. Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence---in the top left corner we see the US, the UK, Germany, Canada, and Israel. These states cluster together in the outer ring of this visualization because they tend to send conflicts to similar targets.\footnote{For details on how to interpret this type of visualization see \citet{minhas:etal:2016:arxiv}.} Conversely, in the bottom right of the outer ring, we see a cluster of authoritarian countries: Iraq, Russia, Syria, North Korea, and China. In the inner ring, the proximity of countries is determined by the degree to which they receive conflict from the same countries. In general, the clusters found on both the inner and outer rings have similar governmental types (Iraq, Syria, Libya, and North Korea all fell under the ``boss" category). In the GLM, which ignores these third-order dependencies, many of these results might have been attributed to regime type. Structural equivalence is present even when accounting for nodal characteristics like regime type.  The AME model, on the other hand, shows that this can be specified in terms of the interdependencies captured by the multiplicative effects.

@@

Fourth, it is clear that the AME approach dominates the GLM approaches in terms of performance. Not only it is better at correctly identifying cases in which the dependent variable takes a value of $0$ (via the ROC curves and associated statistics), but it also dominates at correctly identifying occurrences of the dependent variable in the data (seen via the PR curves and associated statistics).  In the case of studies with continuous dependent variables, the AME approach has average error statistics that are about one-half that found in the OLS model.

>>>>

An important takeaway here is that many scholars are forced to make knowledge claims based on the statistical significance of a small set of covariates, or the differences between these covariates. These differences may change dramatically when interdependencies are taken into account directly. This outcome follows from AME's ability to better account for the dependencies discussed in the previous section, whereas GLM approaches explicitly assume observational independence conditional on the specified covariates. As this is a widely-known limitation of GLM approaches, scholars often attempt to account for clustering of observations by including additional variables and adjusting the standard errors of the resulting estimates. At best, this method introduces noise and imprecision into results, and at worst can produce misleading outcomes.

Beyond just comparing parameter estimates, we examine how well each approach can represent the data generating process using an out-of-sample cross validation strategy. Specifically, for each study, we randomly divide the data into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.

@@

The result of this procedure is a set of sociomatrices $\bm \hat Y$, in which each entry $\hat y_{ij,t}$ is a predicted value obtained from using a subset of the data that does not include $y_{ij,t}$. We summarize the performance of the various models in Table~\ref{tab:modelPerfSumm} below. For the binary models we provide the area under the Receiver Operator Characteristic (ROC) and Precision Recall (PR) curves. Only one of the studies here had a continuous dependent variable and for this we provide the root mean squared error (RMSE) and root median squared error (RMDSE).\footnote{More details on the performance of each of these models can be found in the Appendix.} For each of the replications, we find that the AME approach substantially outperforms the original models in terms of out-of-sample predictive performance. This is important as it indicates that switching to the AME framework---even when using the exact same specification as the original studies---enables scholars to better represent the data generating process of their events of interest. The fact that this analysis is done in an out-of-sample context ensures that the AME framework is not simply overfitting with more parameters, rather the dependence parameters we include are capturing underlying structure previously missed by the exogenous covariates in the models.

@@

In the original model, the variable ``Pers/Democ Directed Dyad" (which represents a Personalist $\rightarrow$ Democractic directed dyad) has a positive effect while the variable ``Democ/Personalist Directed Dyad'' is too imprecisely measured to indicate a direction. In our re-estimation using the AME framework, we also find that Pers-Democ directed dyad has a positive effect while Democ-Pers directed dyad is still too imprecisely measured to indicate a direction. Using this model, however, we can no longer conclusively say that the Pers/Democratic coefficient is larger than the Democ/Personalist one.\footnote{Comparing } Our re-estimation using the AME approach therefore cast some doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies, though many of the major effects  uncovered by the authors remains robust when accounting for dyadic interdependencies. %Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework suggesting that some of their effects were actually the products of interdependencies.

@@

\citet{mcdonald:2004} studies whether trade promotes peace between nations. \citet[p. 547]{mcdonald:2004} introduced the argument that free trade between states ``makes conflict less likely because of its efficiency over conquest in acquiring resources\ldots''. Accordingly, he provided evidence challenging the generalized linkage between peace and trade and refined the measurement of the standard ``trade'' variable, arguing that \textit{free} trade, rather than trade alone, reduces the likelihood of conflict between states. His key hypothesis is that greater levels of protection increase the probability of interstate conflict, an argument that builds on the work of classic liberalism and connects free trade to the power of domestic audiences. \citet{mcdonald:2004} measured free trade in two ways. The first, the ``protection'' variable measures the proportion of customs revenue divided by total imports in the state that possesses the greater such ratio in each dyad. reflects the notion that larger protected sectors generate greater societal pressures resulting in pockets of support for war. This measure thus captures the score of the state in the dyad that possesses higher barriers to trade. \citet[p. 560]{mcdonald:2004} also includes a measure of economic integration  calculated as ``the lower proportion of total dyadic trade (imports plus exports) divided by state $i$'s GDP or total dyadic trade divided by state $j$'s GDP''. The (binary) dependent variable is the onset of a new militarized interstate dispute within a given dyad, and \citet{mcdonald:2004} includes splines to correct for temporal dependence with robust standard errors clustered on each dyad.

@@

When we compare the results of Rose's original OLS model to the AME model accounting for network dependencies, the results are similar. The main result of the model---the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables---does not prevail in the AME model. However, in a sense the AME result provides even stronger evidence against a positive effect of WTO membership: in the original model, this negative linkage is not significant, but in the AME estimation this negative linkage is more precisely estimated. The fact that this variable attains conventional statistical significance under an AME estimation shows an additional value of this approach to applied researchers -- while accounting for interdependencies and removing that source of bias will remove false positive findings, it can also avoid false negative findings where the bias prevented a researcher from uncovering a consistent relationship.\footnote{This benefit comes up less often in these replications because most published studies are positive, rather than null findings.} In both cases we find  a rejection of the conventional wisdom, that WTO membership increases trade. Of course, the controversial result of the paper, strengthened in the AME analysis might be an artifact of measurement error. \citep{tomz:etal:2007}. While in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.

\begin{figure}
	\includegraphics[width=\textwidth]{rose_aeff_top10.pdf}
	\caption{Nodal Random Effects for AME estimation of Rose (2004), for the countries with the highest and lowest Random Effects.}
	\label{fig:roser}
\end{figure}
\FloatBarrier

@@

\citet{weeks:2012} examines the influence of domestic institutions on the initiation of military conflicts by autocratic leaders.  She argues that in some circumstances autocrats are held accountable for their foreign policy decisions. She adds the nuance that autocratic audiences are not homogeneous. When the autocratic regime is nonmilitary, domestic audiences do not favor military actions, but in military autocracies this is not the case. Further she argues that in personalistic regimes without a military or civilian domestic audience, the leaders tend to be more likely to employ military force in their foreign policy.  To study this question, she uses a dyadic design in which the dependent variable is ``whether country A in a directed dyad initiated military conflict against country B during year t'' (page 337).  One major innovation in her study resides in the nuanced way in which she conceptualizes and codes regimes into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen. She also includes a variety of putative control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.'' She uses a logistic regression, but follows \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with fixed, unit effects. The analysis is done for dyads, but is considered to be from the perspective of the actor that initiated the dispute. The basic finding is that a) juntas, boss type, and strongmen type regimes are more likely to initiate conflict than machine-type regimes (and maybe democracies) and that b) machine-type regimes are no more belligerent than democracies.   These insights are mainly emphasized in the paper by the parameter estimates depicted in Tables 1 and 2 (pages 339--340) from the paper. She argued that ignoring important nuances between different types of autocracies hinders our understanding of the initiation of military conflict by autocracies.

\begin{figure}[!h]
	\includegraphics[width=\textwidth]{weeks_circPlot.png}
	\caption{ Visualization of multiplicative effects for Weeks (2012). Each circle designates a country and the color corresponds to the legend at the center of the visualization. Countries that cluster together in the outer ring are those that were found by the model to have similar sending patterns, meaning that they tend to send conflict to similar sets of countries. The inner ring clusters countries by the similarity of who they receive conflict from. 
	%Blue represents groups with common sending patterns and red represents groups with common receiving patterns.
	}
	\label{fig:weekscirc}
\end{figure}
\FloatBarrier

The re-estimation of \citet{weeks:2012} has the sharpest divergence between the GLM results and those of the AME Model. In Weeks' initial models,  she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses, and Strong-men are more conflict-prone. When we examine the AME results, we find that none of these values are distinguishable from zero. Similarly, we find less pronounced effects for military capabilities. One explanation for this divergence is the AME model's ability to account for third-order effects. Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence---in the top left corner we see the US, the UK, Germany, Canada, and Israel. These states cluster together in the outer ring of this visualization because they tend to send conflicts to similar targets. Conversely, in the bottom right of the outer ring, we see a cluster of authoritarian countries: Iraq, Russia, Syria, North Korea, and China. In the inner ring, the proximity of countries is determined by the degree to which they receive conflict from the same countries. In general, the clusters found on both the inner and outer rings have similar governmental types (Iraq, Syria, Libya, and North Korea all fell under the ``boss" category). They tend to initiative and be targets of conflict by similar actors but they are not more likely to initiate

Thus she is kinda right but wrong in that they are more likely to intiate conflict in general ... but they do have a similar pattern of relationshsips

In the GLM, which ignores these third-order dependencies, many of these results might have been attributed to regime type. Structural equivalence is present even when accounting for nodal characteristics like regime type.  The AME model, on the other hand, shows that this can be specified in terms of the interdependencies captured by the multiplicative effects.

@@

Fourth, it is clear that the AME approach dominates the GLM approaches in terms of performance. Not only it is better at correctly identifying cases in which the dependent variable takes a value of $0$ (via the ROC curves and associated statistics), but it also dominates at correctly identifying occurrences of the dependent variable in the data (seen via the PR curves and associated statistics). In the case of the study with a continuous dependent variable, the AME approach has average error statistics that are about one-half that found in the original model.
